---
title: "Create a Title for the Project"
author: "Your Name"
format: 
  dashboard:
    orientation: columns
    embed-resources: true
---

```{r}
#| label: setup
#| include: false

#add any packages that you need here

```

# Introductory Dashboard

Be sure to include:\
- Data source/descriptions;\
- any prerequisite data cleaning;\
- questions to be asked of the data and answered in subsequent visualizations.

# Page 2 - Customize Title!

Default layout (specified in YAML) is COLUMNS, so this text creates a separate column on the dashboard

You can use different formats to include multiple visualizations or just focus on one visualization. Remember to see Canvas for resources! Here is the main website on [Quarto Dashboard Documentation](https://quarto.org/docs/dashboards/).

## Column {width="60%"}

Add text here if desired

### Chart 1. Note that this title doesn't actually show up on the dashboard! The \# is needed to create a title, but the text after it is ignored unless you customize it below.

```{r}
#| title: Add TITLE here

plot(cars)
```

## Column (note I can "title" this heading whatever I want. R knows to make a column because I specified column orientation in the YAML) {width="40%"}

### Chart 2

```{r}
plot(cars)
```

### Chart 3

```{r}
#| title: Chart 3 (customize the title!)

plot(cars)
```

# Page 3 Title {orientation="rows"}

Notice the "row" orientation of this dashboard!

## Row {height="60%"}

### Create interesting Headers

```{r}
plot(cars)
```

## Row {height="40%"}

### Chart 2

```{r}
plot(cars)
```

### Chart 3

```{r}
plot(cars)
```

And remember you can add text...
